\documentclass[10pt]{report}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{url}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{abstract} % Allows abstract customization
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\usepackage{titling}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{bera}	% Font for listings
\lstset{basicstyle=\ttfamily\small,
escapeinside={||},
mathescape=true}
\pagenumbering{arabic}
\usepackage{bm}
\usepackage{upgreek}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage[font=small]{caption}
\usepackage[
%  disable, %turn off todonotes
  colorinlistoftodos, %enable a coloured square in the list of todos
  textwidth=\marginparwidth, %set the width of the todonotes
  textsize=scriptsize, %size of the text in the todonotes
  ]{todonotes}
 \usepackage{subfiles}

\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Somax 2 Developer's Documentation}
\fancyfoot[C]{\thepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{The Somax 2 Developer's Documentation\\[0.24cm]
	\large{Rev. 0.1.0}}
\author{Joakim Borg }
\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\renewcommand{\maketitlehookd}{
%	\begin{abstract}
%	\noindent abstract abstract abstract todo (or maybe skip?)
%	\end{abstract}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle
\chapter*{}
\large{\textbf{Credits}\\}\\
\small{
Somax 2 \copyright \ Ircam 2012-2023\\

\noindent Somax 2 is a renewed version of the Somax reactive co-improvisation paradigm by G. Assayag.
Architecture, UI and code completely redesigned and written by Joakim Borg in Max and Python.
\\

\noindent \textbf{Legacy:}
\begin{itemize}
\item Early Java prototype by Olivier Delerue: adding reactivity to OMax.
\item Versions 0.1 to 1.3 by Laurent Bonnasse-Gahot: conception of the reactive memory and  influence  dimensions model.
\item Versions 1.4 to 1.9  by Axel Chemla-Romeu-Santos: separation of the Python server and object oriented design.
\end{itemize}

\medskip
\noindent The Somax 2 project  is part of the ANR project MERCI (Mixed Musical Reality with Creative Instruments) and the ERC project REACH (Raising Co-creativity in Cyber-Human Musicianship).
\\

\noindent PI : Gérard Assayag\\
Music Representation Team\\
IRCAM STMS Lab (CNRS, Sorbonne University, Ministry of Culture).
\\

\noindent\url{repmus.ircam.fr/impro}
}


\tableofcontents


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%   DOCUMENT   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{overview}
\input{python-architecture}
\input{max-architecture}


\appendix
\input{appendices}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{reference.bib}{}
\bibliographystyle{plain}
\end{document}